---
title: "Built-in varTables"
author: "Dan Kelley (https://orcid.org/0000-0001-7808-5911)"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 72
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{Built-in varTables}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

# Introduction

This vignette discusses two built-in variable-table files.  These may be used
as patterns by users wishing to create their own tables.  Note that the package
requires that the files be in proper YAML format, which can be checked by using
`yaml::yaml.load_file("file.yml")`, where `"file.yml"` is the trial file.  This
function will report an error if the contents are not in proper YAML format.

# The built-in tables

## `adp`

This table is stored in a file that may be accessed with
```{r eval=FALSE}
system.file("extdata", "adp.yml", package="ocencdf")
```
and the contents are as follows.
```{r, echo=FALSE}
#| echo: false
#| fig.width: 10
#| comment: ""
lines <- readLines(system.file("extdata", "adp.yml", package="ocencdf"))
for (line in lines)
    cat(line, "\n")
```


## `adv`

This table is stored in a file that may be accessed with
```{r eval=FALSE}
system.file("extdata", "adv.yml", package="ocencdf")
```
and the contents are as follows.
```{r, echo=FALSE}
#| echo: false
#| fig.width: 10
#| comment: ""
lines <- readLines(system.file("extdata", "adv.yml", package="ocencdf"))
for (line in lines)
    cat(line, "\n")
```


## `argo`

This table, patterned on naming conventions in the Argo ocean float program, is
stored in a file that may be accessed with
```{r eval=FALSE}
system.file("extdata", "argo.yml", package="ocencdf")
```
and the contents are as follows.
```{r, echo=FALSE}
#| echo: false
#| fig.width: 10
#| comment: ""
lines <- readLines(system.file("extdata", "argo.yml", package="ocencdf"))
for (line in lines)
    cat(line, "\n")
```

## `whp`

This table, patterned on naming conventions in the World Hydrographic Program,
is stored in a file that may be accessed with
```{r eval=FALSE}
system.file("extdata", "whp.yml", package="ocencdf")
```
and the contents are as follows.
```{r, echo=FALSE}
#| echo: false
#| fig.width: 10
#| comment: ""
lines <- readLines(system.file("extdata", "whp.yml", package="ocencdf"))
for (line in lines)
    cat(line, "\n")
```

