---
title: "1. Introduction to ocencdf"
author: "Dan Kelley (https://orcid.org/0000-0001-7808-5911)"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
    fig_caption: yes
    fig_width: 3.5
    fig_height: 3.5
    dpi: 72
    dev.args: list(pointsize=11)
vignette: >
  %\VignetteIndexEntry{1. Introduction to ocencdf}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

<!--
Building this vignette.

1. Edit vignettes/oce.Rmd in a plain-text editor.

2. In a unix shell at the level above ocencdf, type the following:

       R CMD build --compact-vignettes="qpdf" ocencdf
       R CMD install ocencdf_0.0-1.tar.gz

   where the version number may need adjustment in the second command.

3. To see the results, type the following in an R console:

      browseVignettes("ocencdf")
-->

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Purpose

The package provides a way to store approximate forms of oce objects as Netcdf
files.  Since Netcdf is handled by many computing languages, it is a good
intermediate format in multi-language processing, as well as a good choice for
data archiving.

# Limitations of Netcdf format

Unfortunately, direct transferral of R information to Netcdf is not possible.
This is because Netcdf is an array-oriented format that is not designed to hold
tree-oriented data such as R lists (which are used throughout oce).  It is
possible to flatten simple lists, but the process gets complicated for nested
lists, which are common in oce.  This is not the only problem with Netcdf. It
also lacks basic elements such as user-defined classes and variable-length
character strings.  Related to this second point is the fact that Netcdf
tolerates only a single unlimited dimension.

This package address these and other Netcdf limitations by focussing effort only
on certain important oce classes, and only certain elements within those
classes.  The initial focus is on ctd data, mainly because they are of greatest
importance in everyday work.

# How the package works

## Focus elements

The design goal is to save the following elements

1. Entries in the `data` slot (e.g. temperature, salinity, etc.)
2. Entries in the `flags` entry of the `metadata` slot, as entries with names ending in `_QC`.
3. An approximate form of units, as strings.
4. The `header` item in the `metadata` slot.
5. A few other entries from the `metadata` slot, e.g. `longitude` and `latitude`.

Note that other `metadata` may be added as needed or requested, but that there
is no intention of trying to copy everything over, mainly because of the
awkardness of saving variable-length character strings and lists in Netcdf.

Once, or if, significant progress is made on ctd data, the next object type
to examine will likely be adcp data.

## Renaming elements

Oce uses a restricted set of names for certain variables.  For example, in ctd
objects, temperature is called `temperature`, and if the ctd had two temperature
sensors, there would be an additional entry called `temperature2`.  These are
not typically the names used in raw data files, however.  More commonly,
temperature might be named `TEMP`, for example.  In recognition of this, ocencdf
provides a way to rename oce objects using other systems of names.  The mapping
between oce name and other name is controlled by YAML (yet another markup
language) files, which are called varTables in this package. Tables are provided
for the Argo system and for the World Hydographic Program system. These are
available as `"argo"` and `"whp"`, respectively.  They are stored in the
`inst/extdata` directory of the package source tree. Users seeking to define new
naming systems must copy the formats of those files.

# Example with CTD data

The first step is to create an oce object, e.g. station 100 of the built-in
`section` dataset.  Note the variable names, units, and flag values in the
summary.

```{r}
library(ocencdf)
library(oce)
data(section)
stn <- section[["station", 100]]
summary(stn)
```

Now, save as a netcdf file, using (say) the Argo convention for variable names.

```{r}
oce2ncdf(stn, varTable="argo", ncfile="stn.nc")
```

The `stn.nc` file may be examined in various ways, but we leave that up to the
reader.  Note that the column names are e.g. `TEMP` in the Netcdf file,
because the varTable named `"argo"` establishes that nickname.

In R, we can read `stn.nc` and get a summary with the following.

```{r}
A <- ncdf2oce("stn.nc", varTable="argo")
summary(A)
```

Note that the variable names have been converted from Netcdf names (e.g. `"TEMP"`) to
oce names (e.g. `"temperature"`).  Units have also been decoded (again, according to
the `"argo"` varTable).  However, notice that the flags are being displayed as though
they were data columns (e.g. `temperature_QC`).  This is because `ncdf2oce()` does
not try to translate these into flags.  To do that, use `as.ctd()`, as follows.

```{r}
B <- as.ctd(A)
summary(B)
```

**Exercise:** Try using `ncdf2ctd()` instead of the two-step procedure used for
`A` and `B`. Do you get the same results?

```{r echo=FALSE}
unlink("stn.nc")
```

